Measuring contrast processing in the visual system using SSVEP

### 2024-04-07 review.qmd
Removed duplicate figure caption
Added plt.show

Epilepsy is a neurological condition in which patients experience seizures - episodes of uncontrolled neural activity that can cause unconsciousness, involuntary movements and convulsions, and atypical sensory experiences. @Porciatti2000 showed that individuals with photosensitive epilepsy generate larger steady-state signals in response to flickering visual stimuli, and their contrast response functions saturate less than those of healthy controls. This is consistent with the idea that epilepsy involves a cortical hyperexciteability that makes seizures more likely. It is also the case for individuals with idiopathic generalised epilepsy [@Tsai2011], a subtype of epilepsy that has a less obvious link to vision. The differences apply across the whole contrast-response function, and so resemble a response gain effect, which might be due to reduced inhibition from neighbouring neurons. Differences in SSVEP amplitudes have also been reported in individuals with migraine [@Shibata2008], a condition also associated with cortical hyperexciteability.

@@

```{python}
#| label: fig-plotclinical
#| fig-cap: "Example contrast response functions for different clinical conditions. Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al. (2011). Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al. (2015). Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al. (2018). Panel (d)..."
#| fig-cap-location: bottom

@@

plt.subplot(2, 2, 4)  # Fourth plot for Parkinson's
plt.plot(finecontrastsdB, respParkPatients, lw=4, color='red', label='Parkinson\'s patients')
plt.plot(finecontrastsdB, respParkControls, lw=4, color='black', linestyle='--', label='Control participants')
plt.xticks(ticklocsx, ticklabelsx)
plt.yticks(np.arange(0, 7, 2))
plt.xlabel('Target contrast (%)')
plt.ylabel('SSVEP response')
plt.title("(d) Parkinson's")
plt.legend()
# To save the figure
# plt.savefig('Figures/clinicalplot.pdf')

@@

![Example contrast response functions for different clinical conditions. Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al. (2011). Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al. (2015). Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al. (2018). Panel (d)...](Figures/clinicalplot.pdf){#fig-plotclinical}

>>>>

Epilepsy is a neurological condition in which patients experience seizures - episodes of uncontrolled neural activity that can cause unconsciousness, involuntary movements and convulsions, and atypical sensory experiences. @Porciatti2000 showed that individuals with photosensitive epilepsy generate larger steady-state signals in response to flickering visual stimuli, and their contrast response functions saturate less than those of healthy controls. This is consistent with the idea that epilepsy involves a cortical hyperexciteability that makes seizures more likely. It is also the case for individuals with idiopathic generalised epilepsy [@Tsai2011], a subtype of epilepsy that has a less obvious link to vision. The differences apply across the whole contrast-response function, and so resemble a response gain effect (see [@fig-plotclinical]a), which might be due to reduced inhibition from neighbouring neurons. Differences in SSVEP amplitudes have also been reported in individuals with migraine [@Shibata2008], a condition also associated with cortical hyperexciteability.

@@

```{python}
#| label: fig-plotclinical
#| fig-cap: "Example contrast response functions for different clinical conditions. Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al. (2011). Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al. (2015). Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al. (2018). Panel (d)"
#| fig-cap-location: bottom

@@

plt.subplot(2, 2, 4)  # Fourth plot for Parkinson's
plt.plot(finecontrastsdB, respParkPatients, lw=4, color='red', label='Parkinson\'s patients')
plt.plot(finecontrastsdB, respParkControls, lw=4, color='black', linestyle='--', label='Control participants')
plt.xticks(ticklocsx, ticklabelsx)
plt.yticks(np.arange(0, 7, 2))
plt.xlabel('Target contrast (%)')
plt.ylabel('SSVEP response')
plt.title("(d) Parkinson's")
plt.legend()
# To save the figure
# plt.savefig('Figures/clinicalplot.pdf')
plt.show()

@@
